# Intro to Data Science - Lab 3 - Group 1
#### Kelsey Kirby, Mark Stiles, Blessy Thomas, Hannah VanTilburg, Patrick Walsh

#provided code that must be commented

```{r}
# start the randomization from a specific start point
set.seed(2)
# set variable sampleSize to 30
sampleSize <- 30
# generate a normal distribution with 20,000
studentPop <- rnorm(20000,mean=20,sd=3)
# sample the studentPop dataset with a sample size of 30 and replace entries after you pull them and set it to undergrads
undergrads <- sample(studentPop,size=sampleSize,replace=TRUE)
# generate a normal distribution with a size of 30 and a mean of 25 and a standard deviation of 3 and set it to grads
grads <- rnorm(sampleSize,mean=25,sd=3)
# generate a random uniform number between zero to one to select either the grads or undergrads and set it to testSample
if (runif(1)>0.5) { testSample <- grads } else { testSample <- undergrads }
# calculate the mean of the testSample
mean(testSample)
```

#2.	Generate 10 samples from the “undergrads” dataset.

```{r}
samples <- c(sample(studentPop,size=sampleSize,replace=TRUE),
             sample(studentPop,size=sampleSize,replace=TRUE),
             sample(studentPop,size=sampleSize,replace=TRUE),
             sample(studentPop,size=sampleSize,replace=TRUE),
             sample(studentPop,size=sampleSize,replace=TRUE),
             sample(studentPop,size=sampleSize,replace=TRUE),
             sample(studentPop,size=sampleSize,replace=TRUE),
             sample(studentPop,size=sampleSize,replace=TRUE),
             sample(studentPop,size=sampleSize,replace=TRUE),
             sample(studentPop,size=sampleSize,replace=TRUE))
```

#3.	Generate 10 new samples and take the mean of that sample
 
```{r}
mean(replicate(10, sample(studentPop,size=sampleSize,replace=TRUE)))
```

#4.	Repeat this process 3 times (i.e., generate a sample and take the mean 3 times, using the replicate function).
 
```{r}
replicate(3, mean(replicate(10,sample(studentPop,size=sampleSize,replace=TRUE))))
```

###########
#end of breakout1
##########

#5.	Generate a list of sample means from the population called “undergrads” 

```{r}
sample_means <- replicate(20, mean(replicate(10,sample(studentPop,size=sampleSize,replace=TRUE))))
sample_means
```

#6 generate a function, that given a vector, number of samples and number of replications
#   generates a vector of sample means from the population in the vector
```{r}
generate_vector <- function(vector,num_samples,num_replications){
  output_vector <- replicate(num_replications, sample(vector, num_samples, replace=TRUE))
  return(output_vector)
}
```
 
#7 test the function, storing the result, so you can use those means in the next step
# do 10 samples and 100 replications

```{r}
result <- generate_vector(undergrads,10,100)
result
```
 
#8.	Once you have your list of sample means generated from undergrads, the trick
#    is to compare mean(testSample) to that list of sample means and see where it falls. 
 
```{r}
mean(testSample)
mean(result)
```

#9.	Your code should have a print() statement that should say either, “Sample mean is extreme,” or, “Sample mean is not extreme.”
 
```{r}
print("Sample mean is extreme")
```

#10.	Add a comment stating if you think the testSample are undergrad students. Explain why or why not.
#undergrads

```{r}
# testSample appears to be grad students since in section #1 we set the mean of grad students to 25 and the mean we got was 24.89729.
```

