# Intro to Data Science - Lab 4 - Group 1
#### Kelsey Kirby, Mark Stiles, Blessy Thomas, Hannah VanTilburg, Patrick Walsh

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Get a frame of data from the CitiBike NYC project using JSON
```{r}
library(RCurl)   # For access to Internet data
library(jsonlite) # For decoding JSON
```

```{r}
station_link <- 'https://gbfs.citibikenyc.com/gbfs/en/station_status.json'
apiOutput <- getURL(station_link) 
#prettify(apiOutput)
apiData <- fromJSON(apiOutput) 
stationStatus <- apiData$data$stations
cols <- c('num_bikes_disabled','num_docks_disabled', 'station_id',
          'num_ebikes_available', 'num_bikes_available', 'num_docks_available')
stationStatus = stationStatus[,cols]
#stationStatus
```

#1. Explain what you see if you type in the station_link URL into a browser (in a comment, write what you see)
```{r}
#I see text with a lot of repeating values like station_status, station_status, legacy_id, is_installed etc. It's JSON code that represents the citi bike station information. 
```

#2. Provide a comment explaining each line of code. 
```{r}
# This line stores a URL in a variable
#station_link <- 'https://gbfs.citibikenyc.com/gbfs/en/station_status.json'

# This line uses the url to get the data from the URl 
#apiOutput <- getURL(station_link) 

# This line writes to the console the values in a heirarchical structure
#prettify(apiOutput)

# This line converts the string to a JSON object in R
#apiData <- fromJSON(apiOutput) 

# This line drills down into the object to get the station information, skipping the rest of the request data
#stationStatus <- apiData$data$stations

# This line creates a new vector of the names of variables in the station data
#cols <- c('num_bikes_disabled','num_docks_disabled', 'station_id', 'num_ebikes_available', 'num_bikes_available', 'num_docks_available')

# This line subsets the station data by getting all rows but only the 6 variables from the previous vector
#stationStatus = stationStatus[,cols]
```

#3. Use str( ) to find out the structure of apiOutput and apiData. Report (via a comment) what you found and explain the difference between these two objects. 

```{r}
# This is a string of the whole value of the JSON object
str(apiOutput <- getURL(station_link))
```

```{r}
# This is a JSON object which consists of a List of 2 ints and another List
str(apiData <- fromJSON(apiOutput))
```
 
#4. The apiOutput object can also be examined with a custom function from the jsonlite package called prettify( ). Run this command and explain what you found (in a comment). 
```{r}
# Prettify adds indentation to a JSON string making it easier to read
prettify(apiOutput)
```

#5. Explain stationStatus (what type of object, what information is available)
```{r}
# stationStatus is a data frame. It contains information (number of bikes disabled, station id, number of docks available etc.) about any single bike station
str(stationStatus)
```

#6. Generate a histogram of the number of docks available 
```{r}
hist(stationStatus$num_docks_available)
```
 
#7. Generate a histogram of the number of bikes available 
```{r}
hist(stationStatus$num_bikes_available)
```

#8. How many stations have at least one ebike?
```{r}
numEBikes <- nrow(stationStatus[stationStatus$num_ebikes_available > 1,])
paste("There are", numEBikes, "of stations with at least 1 eBike")
``` 
 
#9. Explore stations with at least one ebike by creating a new dataframe, that only has stations with at least one eBike. 
```{r}
ebikeDF <- stationStatus[stationStatus$num_ebikes_available > 1,]
str(ebikeDF)
hist(ebikeDF$num_ebikes_available)
hist(ebikeDF$num_bikes_available)
hist(ebikeDF$num_docks_available)
```

#10. Calculate the mean of ‘num_docks_available’ for this new dataframe. 
```{r}
docksMean <- mean(ebikeDF$num_docks_available)
paste("The mean number of docks available is", docksMean)
```
 
#11. Calculate the mean of ‘num_docks_available’ for for the full ‘stationStatus’ dataframe. In a comment, explain how different are the two means? 
```{r}
allDocksMean <- mean(stationStatus$num_docks_available)
paste("The mean number of docks available for all stations is", allDocksMean)
```

#12. Create a new attribute, called ‘stationSize’, which is the total number of “slots” available for a bike (that might, or might not, have a bike in it now). Run a histogram on this variable and review the distribution.
```{r}
stationStatus$stationSize <- stationStatus$num_ebikes_available + stationStatus$num_bikes_available + stationStatus$num_docks_available
hist(stationStatus$stationSize)
```
 
#13. Use the plot( ) command to produce an X-Y scatter plot with the number of occupied docks on the X-axis and the number of available bikes on the Y-axis. Explain the results plot.
```{r}
library(ggplot2)
# This is the plot of 'unoccupied docks' vs 'available bikes'.There is a negative correlation between empty docks and available bikes because they are directly related. If a bike is available then one less dock is unoccupied and vice versa. 
ggplot(stationStatus) +
  aes(x=num_docks_available, y=(num_ebikes_available + num_bikes_available)) +
  geom_point() +
  xlab("Number of Docks Available") +
  ylab("Number of Bikes Available")
```
```{r}
# This is the number of 'occupied docks' (available bikes) vs available bikes. (Not sure if this was a typo). There is a positive correlation between bikes being available and bikes being available
ggplot(stationStatus) +
  aes(x=(num_ebikes_available + num_bikes_available), y=(num_ebikes_available + num_bikes_available)) +
  geom_point() +
  xlab("Number of Docks Available") +
  ylab("Number of Bikes Available")
```


