# Introduction to Data Science HW 4
##### Copyright Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva


```{r}
# Enter your name here: Patrick Walsh
```

### Attribution statement: (choose only one and delete the rest)


```{r}
# 1. I did this homework by myself, with help from the book and the professor.
```

Reminders of things to practice from previous weeks: <br>
Descriptive statistics: 	mean( ) max( ) min( )<br>
Coerce to numeric:		as.numeric( )<br>

## Part 1: Use the Starter Code 

Below, I have provided a starter file to help you. <br>

Each of these lines of code **must be commented** (the comment must that explains what is going on, so that I know you understand the code and results).


```{r}
library(jsonlite)  # load the jsonlite library
dataset <- url("https://intro-datascience.s3.us-east-2.amazonaws.com/role.json")  # retrieve this JSON dataset from the URL and save it to variable called 'dataset'
readlines <- jsonlite::fromJSON(dataset)  # converts JSON object to R object  
df <- readlines$objects$person  # saves readlines objects as a dataframe called 'df'
```

A.	Explore the **df** dataframe (e.g., using head() or whatever you think is best).


```{r}
library(tidyverse)
glimpse(df)

```

B.	Explain the dataset <br>
o	What is the dataset about? <br>
o	How many rows are there and what does a row represent?<br>
o	How many columns and what does each column represent?<br>



```{r}
# The dataset contains data about U.S. senators. There are 100 senators at any given time serving in the U.S. Senate.
# There are 100 rows and each row represents a different senator.
# There are 17 columns and each column represents an attribute or feature of that senator, such as their name, gender, website, and twitter id.

```

C. What does running this line of code do? Explain in a comment:


```{r}
# Uses the substr() function which extracts a portion of the birthday column from the df dataframe and saves it
# to a variable called 'vals'. It slices out the first 4 characters only, giving us the year of the birtdate.
vals <- substr(df$birthday,1,4)
vals
```

D. Create a new attribute 'age' - how old the person is 
**Hint:** You may need to convert it to numeric first.


```{r}
# I am using the eeptools library to accomplish this task. I first grab today's date using the Sys.Date()
# function and calculate the age of the senators by using the age_calc() function using today's date and
# the birthdate, displaying the output in years.


library("eeptools")  # load library into RStudio

date_today <-  Sys.Date()  # get today's date
date_today

birthdate <- as.Date(df$birthday)  # save vector of birthdates from dataframe to variable called 'birthdate'
birthdate

age <- age_calc(birthdate, date_today, unit="years")  # saves vector of ages to variable called 'age'
age
```

E. Create a function that reads in the role json dataset, and adds the age attribute to the dataframe, and returns that dataframe


```{r}
add_age <- function(age){
  df$age = age  # add a vector to the dataframe
  return(df)
}

```

F. Use (call, invoke) the function, and store the results in df


```{r}
df <- add_age(age)
glimpse(df)
```

## Part 2: Investigate the resulting dataframe 'df'

A.	How many senators are women? 


```{r}
length(which(df$gender=="female"))
```

B.	How many senators have a YouTube account?  


```{r}
length(which(!is.na(df$youtubeid)))
```

C.	 How many women senators have a YouTube account?


```{r}

nrow(df %>% filter(gender == "female", !is.na(youtubeid)))
```

D.	Create a new dataframe called **youtubeWomen** that only includes women senators who have a YouTube account.   


```{r}
youtubewomen <- df %>% filter(gender == "female", !is.na(youtubeid))
glimpse(youtubewomen)
```

E.	Make a histogram of the **age** of senators in **youtubeWomen**, and then another for the senators in **df**. Add a comment describing the shape of the distributions.


```{r}
hist(youtubewomen$age)
```


```{r}
hist(df$age)
```
```{r}
# The youtubewomen$age histogram is left-skewed whereas the df$age histogram is fairly normally distrubuted.
# This means that the mean and mode in df$age histogram are close to the same, while in the the youtubewomen$age
# histogram, the mean is to the left of the mode.
```

