# Intro to Data Science - HW 5
##### Copyright Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva

```{r}
# Enter your name here: Patrick Walsh
```


### Attribution statement: (choose only one and delete the rest)


```{r}
# 1. I did this homework by myself, with help from the book and the professor.
```

**This module:** **Data visualization** is important because many people can make sense of data more easily when it is presented in graphic form. As a data scientist, you will have to present complex data to decision makers in a form that makes the data interpretable for them. From your experience with Excel and other tools, you know that there are a variety of **common data visualizations** (e.g., pie charts). How many of them can you name? <br>

The most powerful tool for data visualization in R is called **ggplot**. Written by computer/data scientist **Hadley Wickham**, this **“graphics grammar”** tool builds visualizations in layers. This method provides immense flexibility, but takes a bit of practice to master.


## Step 1: Make a copy of the data

A.	Read the **who** dataset from this URL: https://intro-datascience.s3.us-east-2.amazonaws.com/who.csv <br>
into a new dataframe called **tb**. 

Your new dataframe, tb, contains a so-called **multivariate time series**: a sequence of measurements on 23 Tuberculosis-related (TB) variables captured repeatedly over time (1980-2013). Familiarize yourself with the nature of the 23 variables by consulting the dataset’s codebook which can be found here: https://intro-datascience.s3.us-east-2.amazonaws.com/TB_data_dictionary_2021-02-06.csv.


```{r}
library(tidyverse)
tb <- read.csv("https://intro-datascience.s3.us-east-2.amazonaws.com/who.csv")
glimpse(tb)
```

B.	How often were these measurements taken (in other words, at what frequency were the variables measured)? Put your answer in a comment. 


```{r}
# Each variable was taken once a year for each iso2
```

## Step 2: Clean-up the NAs and create a subset

A.	Let's clean up the iso2 attribute in **tb** 

Hint: use *is.na()* -- well use *! is.na()* 


```{r}
tb_subset <- subset(tb, !is.na(iso2))
sum(is.na(tb_subset$iso2))  # verify that there are no NA values left in iso2 column
```

B. Create a subset of **tb** containing **only the records for Canada ("CA" in the iso2 variable)**. Save it in a new dataframe called **tbCan**. Make sure this new df has **29 observations and 23 variables**.


```{r}
tbCan <- subset(tb_subset, iso2=="CA")
glimpse(tbCan)
```

C. A simple method for dealing with small amounts of **missing data** in a numeric variable is to **substitute the mean of the variable in place of each missing datum**. <br>
This expression locates (and reports to the console) all the missing data elements in the variable measuring the **number of positive pulmonary smear tests for male children 0-4 years old** (there are 26 data points missing)


```{r}
tbCan$new_sp_m04[is.na(tbCan$new_sp_m04)]
```


    Error in eval(expr, envir, enclos): object 'tbCan' not found
    Traceback:



D.	Write a comment describing how that statement works.


```{r}
# The first part of the code reads the new_sp_m04 column from the tbCan dataframe: tbCan$new_sp_m04.
# Next, square brackets are used to index certain elements from that dataframe and column.
# Within the square brackets, we use the is.na() function to select only the elements of that dataframe
# and column that are NA.
```

E.	Write 4 more statements to check if there is missing data for the number of positive pulmonary smear tests for: **male and female** children 0-14 years old (**new_sp_m014** and **new_sp_f014**), and **male and female citizens 65 years of age and older**, respectively. What does empty output suggest about the number of missing observations?


```{r}
tbCan$new_sp_m014[is.na(tbCan$new_sp_m014)]
tbCan$new_sp_f014[is.na(tbCan$new_sp_f014)]
tbCan$new_sp_m65[is.na(tbCan$new_sp_m65)]
tbCan$new_sp_f65[is.na(tbCan$new_sp_f65)]

# Empty output means that there are no NA values in that column.
```

There is an R package called **imputeTS** specifically designed to repair missing values in time series data. We will use this instead of mean substitution.  <br>
The **na_interpolation()** function in this package takes advantage of a unique characteristic of time series data: **neighboring points in time can be used to “guess” about a missing value in between**.

F.	Install the **imputeTS** package (if needed) and use **na_interpolation( )** on the variable from part C. Don’t forget that you need to save the results back to the **tbCan** dataframe. Also update any attribute discussed in part E (if needed).


```{r}
#install.packages("imputeTS")
library(imputeTS)
tbCan$new_sp_m04 <- na_interpolation(tbCan$new_sp_m04)
```

G.	Rerun the code from C and E above to check that all missing data have been fixed. 


```{r}
# From Step C:
tbCan$new_sp_m04[is.na(tbCan$new_sp_m04)]

# From Step E:
tbCan$new_sp_m014[is.na(tbCan$new_sp_m014)]
tbCan$new_sp_f014[is.na(tbCan$new_sp_f014)]
tbCan$new_sp_m65[is.na(tbCan$new_sp_m65)]
tbCan$new_sp_f65[is.na(tbCan$new_sp_f65)]
```

## Step 3: Use ggplot to explore the distribution of each variable
**Don’t forget to install and library the ggplot2 package.** Then: <br>
H.	Create a histogram for **new_sp_m014**. Be sure to add a title and briefly describe what the histogram means in a comment.


```{r}
install.packages("ggplot2")
library(ggplot2)

ggplot(tbCan, aes(x=new_sp_m014))+
  ggtitle("New pulmonary smear positive cases: males aged 0-14 years")+
  labs(x="Positive cases", y="Count of positive cases")+
  geom_histogram()

# Histogram shows the positive cases along the X  axis and number of positive cases along the Y axis.
```

I.	Create histograms (using ggplot) of each of the other three variables from E with ggplot( ). <br>
Which parameter do you need to adjust to make the other histograms look right?


```{r}
# You need to adjust the binwidth parameter so that the columns shown in the histograms fit the data. 
# You should also adjust the title to match the data being presented.

ggplot(tbCan, aes(x=new_sp_f014))+
  ggtitle("New pulmonary smear positive cases: females aged 0-14 years")+
  labs(x="Positive cases", y="Count of positive cases")+
  geom_histogram(binwidth = 0.7)
```


```{r}
ggplot(tbCan, aes(x=new_sp_m65))+
  ggtitle("New pulmonary smear positive cases: males aged 65 and over")+
  labs(x="Positive cases", y="Count of positive cases")+
  geom_histogram(binwidth = 3)
```


```{r}
ggplot(tbCan, aes(x=new_sp_f65))+
  ggtitle("New pulmonary smear positive cases: females aged 65 and over")+
  labs(x="Positive cases", y="Count of positive cases")+
  geom_histogram(binwidth = 2)
```

## Step 4: Explore how the data changes over time
J.	These data were collected in a period of several decades (1980-2013). You can thus observe changes over time with the help of a line chart. Create a **line chart**, with **year** on the X-axis and **new_sp_m014** on the Y-axis.


```{r}
ggplot(tbCan, aes(x=year, y=new_sp_m014))+
  ggtitle("New pulmonary smear positive cases: males aged 0-14")+
  labs(x="Positive cases", y="Count of positive cases")+
  geom_line()
```

K.	Next, create similar graphs for each of the other three variables. Change the **color** of the line plots (any color you want).


```{r}
ggplot(tbCan, aes(x=year, y=new_sp_f014))+
  ggtitle("New pulmonary smear positive cases: females aged 0-14")+
  labs(x="Positive cases", y="Count of positive cases")+
  geom_line(color="red")
```


```{r}
ggplot(tbCan, aes(x=year, y=new_sp_m65))+
  ggtitle("New pulmonary smear positive cases: males aged 65 and older")+
  labs(x="Positive cases", y="Count of positive cases")+
  geom_line(color="blue")
```


```{r}
ggplot(tbCan, aes(x=year, y=new_sp_f65))+
  ggtitle("New pulmonary smear positive cases: females aged 65 and older")+
  labs(x="Positive cases", y="Count of positive cases")+
  geom_line(color="green")
```

L.	Using vector math, create a new variable by combining the numbers from **new_sp_m014** and **new_sp_f014**. Save the resulting vector as a new variable in the **tbCan** df called **new_sp_combined014**. This new variable represents the number of positive pulmonary smear tests for male AND female children between the ages of 0 and 14 years of age. Do the same for SP **tests among citizens 65 years of age and older** and save the resulting vector in the tbCan variable called **new_sp_combined65**.


```{r}
# Create new vectors that combine the values of two other vectors
new_sp_combined014 <- c(tbCan$new_sp_m014 + tbCan$new_sp_f014)
new_sp_combined65 <- c(tbCan$new_sp_m65 + tbCan$new_sp_f65)

# Add these new vectors to the dataframe
tbCan <- cbind(tbCan, new_sp_combined014, new_sp_combined65)
glimpse(tbCan)
```

M.	Finally, create a **scatter plot**, showing **new_sp_combined014** on the x axis, **new_sp_combined65** on the y axis, and having the **color and size** of the point represent **year**.


```{r}
ggplot(tbCan, aes(x=new_sp_combined014, y=new_sp_combined65))+
  ggtitle("New pulmonary smear positive cases: 0-14 & 65 and over")+
  labs(x="Cases aged 0-14", y="Cases aged 65 and over")+
  geom_point(aes(color=year, size=year))
```

N.	Interpret this visualization – what insight does it provide?


```{r}
# It shows that the number of positive TB cases have been in decline from 1980 to 2000.
# The scatter plot uses both color and size to show that the concentration of data
# increases as the size of the points get larger (towards 2000) and turn from black to blue (towards 2000).
```
