---
title: "Intro to Data Science - HW 10 - Shiny"
output: html_notebook
runtime: shiny
---

##### Copyright Jeffrey Stanton, Jeffrey Saltz, Christopher Dunham, and Jasmina Tacheva

```{r}
# Enter your name here: Patrick Walsh
```

### Attribution statement: (choose only one and delete the rest)

```{r}
# 1. I did this homework by myself, with help from the book and the professor.
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r}
#install.packages("shiny")
library(tidyverse)
library(shiny)
```

At a high level, we want to create a scatter plot, where the user can select what is the x axis and the y axis of the scatter plot. We will use the same dataset as used in our first ggplot homework.

The size and color will be represented by the year attribute, and the user will be able to choose one of the following (for  the x and the y axis):
new_sp_m014
new_sp_f014
new_sp_m65
new_sp_f65

## Read in the same dataset we used for the ggplot homework

The file is: "https://intro-datascience.s3.us-east-2.amazonaws.com/who.csv" and store it in the *tb* dataframe
```{r}
tb <- read_csv("https://intro-datascience.s3.us-east-2.amazonaws.com/who.csv")
```
## Clean up the dataset, just like we did in the ggplot homework.

First, remove na's from iso2
```{r}
isoNotNA <- !is.na(tb$iso2)
tb <- tb %>% filter(isoNotNA)
glimpse(tb)
```

Now create the dataframe 'tbCan', which is the iso2 for canada (CA)
```{r}
isoIsCA <- tb$iso2=="CA"
tbCan <- tb[isoIsCA,]
glimpse(tbCan)
```

We will need the imputeTS package (only install if needed)
```{r}
#install.packages('imputeTS')
library(imputeTS)
```

Now we can use 'na_interpolation for new_sp_m014
```{r}
tbCan$new_sp_m04 <- na_interpolation(tbCan$new_sp_m04)
```

## Create the User Interface of our shiny app

```{r}
# Define UI for application (called ui)
# Use a 'fluidPage' using the sidebarLayout, and your sidePanel and mainPanel

choices <- c("new_sp_m014", "new_sp_f014", "new_sp_m65", "new_sp_f65")

ui <- fluidPage(

  # Application title
  titlePanel("Exploring Data"),
  
  # Using sidebarLayout 
  sidebarLayout(

      # Define the sidebarPanel, which has two dropdown boxes (use 'selectInput'), 
      # one for the x-axis and the other is the y-axis for the scatter plot
      sidebarPanel(
        selectInput(
          "yaxis", 
          "Y-Axis", 
          choices,
          selected = NULL,
          multiple = FALSE,
          selectize = TRUE,
          width = NULL,
          size = NULL),
        selectInput(
          "xaxis", 
          "X-Axis", 
          choices,
          selected = NULL,
          multiple = FALSE,
          selectize = TRUE,
          width = NULL,
          size = NULL)
        
      ),
      
      # Create the mainPanel to show the scatter plot
      mainPanel(
        plotOutput("scatterPlot")
      )
   )
)
```

## Now let's define the server part of the Shiny app

```{r}
server <- function(input, output) 
{
  
  # output the plot with renderPlot()
  
  # inside the renderPlot() function you would want to:
     ## Use the input from the 2 selectInput() created in the UI portion to filter data from the tbCan dataframe
     ## and make the scatteplot with ggplot2. The scatterplot will be automatically updated whenever there is 
     ## change in one of the selectInput()
  
     
    output$scatterPlot <- renderPlot({
      xvals <- unlist(tbCan[,input$xaxis])
      yvals <- unlist(tbCan[,input$yaxis])
      xlabel <- as.character(input$xaxis)
      ylabel <- as.character(input$yaxis)
      ggplot(tbCan) + 
        aes(x=xvals, y=yvals, color=year, size=year) +
        xlab(xlabel) +
        ylab(ylabel) + 
        geom_point()
    })
}

```

## Now run the shiny App
```{r}
# Run the application 
shinyApp(ui = ui, server = server)
```
