# Intro to Data Science HW 7
##### Copyright Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva


```{r}
# Enter your name here: Patrick Walsh
```

### Attribution statement: (choose only one and delete the rest)


```{r}
# 1. I did this homework by myself, with help from the book and the professor.
```

The chapter on **linear models** (“Lining Up Our Models”) introduces **linear predictive modeling** using the tool known as **multiple regression**. The term “multiple regression” has an odd history, dating back to an early scientific observation of a phenomenon called **“regression to the mean.”** These days, multiple regression is just an interesting name for using **linear modeling** to assess the **connection between one or more predictor variables and an outcome variable**. 


<br>In this exercise, you will **predict Ozone air levels from three predictors**.

A.	We will be using the **airquality** data set available in R. Copy it into a dataframe called **air** and use the appropriate functions to **summarize the data**. 


```{r}
library(tidyverse)
air <- data.frame(airquality)
glimpse(air)
```

B.	In the analysis that follows, **Ozone** will be considered as the **outcome variable**, and **Solar.R**, **Wind**, and **Temp** as the **predictors**. Add a comment to briefly explain the outcome and predictor variables in the dataframe using **?airquality**.


```{r}
# Ozone: Mean ozone in parts per billion
# Solar.R: Solar radiation in Langleys (frequency band)
# Wind: Average wind speed in miles per hour
# Temp: Maximum daily temperature in degrees Fahrenheit
?airquality
```

C.	Inspect the outcome and predictor variables – are there any missing values? Show the code you used to check for that.


```{r}
# The outcome variable Ozone has missing values
# The predictor variable Solar.R has missing values
anyNA(air$Ozone)
anyNA(air$Solar.R)
anyNA(air$Wind)
anyNA(air$Temp)
```

D.	Use the **na_interpolation()** function from the **imputeTS package** (remember this was used in a previous HW) to fill in the missing values in each of the 4 columns. Make sure there are no more missing values using the commands from Step C.


```{r}
#install.packages("imputeTS")
library(imputeTS)
air$Ozone <- na_interpolation(air$Ozone)
air$Solar.R <- na_interpolation(air$Solar.R)

anyNA(air$Ozone)
anyNA(air$Solar.R)
anyNA(air$Wind)
anyNA(air$Temp)
```

E.	Create **3 bivariate scatterplots (X-Y) plots** (using ggplot), for each of the predictors with the outcome. **Hint:** In each case, put **Ozone on the Y-axis**, and a **predictor on the X-axis**. Add a comment to each, describing the plot and explaining whether there appears to be a **linear relationship** between the outcome variable and the respective predictor.


```{r}
library(ggplot2)
```


```{r}
# There appears to be a linear relationships between Solar.R and Ozone where
# Ozone increases a tiny amount as Solar.R increases (positive correlation)
ggplot(air, aes(x=Solar.R, y=Ozone)) +
    geom_point()
```


```{r}
# There appears to be a linear relationship between Wind and Ozone where
# Ozone decreases as Wind increases (negative correlation)
ggplot(air, aes(x=Wind, y=Ozone)) +
    geom_point()
```


```{r}
# There appears to be a linear relationships between Temp and Ozone where
# Ozone increases as Temp increases (positive correlation)
ggplot(air, aes(x=Temp, y=Ozone)) +
    geom_point()
```

F.	Next, create a **simple regression model** predicting **Ozone based on Wind**, using the **lm( )** command. In a comment, report the **coefficient** (aka **slope** or **beta weight**) of **Wind** in the regression output and, **if it is statistically significant**, **interpret it** with respect to **Ozone**. Report the **adjusted R-squared** of the model and try to explain what it means. 


```{r}
# The coefficient of Wind is -4.5925. The p-value is less than 0.05, so it is statistically significant.
# The coefficient of Wind means that as Wind increases by 1, the value for Ozone decreases by -4.5925.
# This shows a negative correlation between Wind and Ozone.
# The Adjusted R-squared is 0.2527, which means that ~25.3% of the change in Ozone can be explained by a
# change in Wind.
lm_simple_regression <- lm(data=air, Ozone ~ Wind)
summary(lm_simple_regression)
```

G.	Create a **multiple regression model** predicting **Ozone** based on **Solar.R**, **Wind**, and **Temp**.<br> **Make sure to include all three predictors in one model – NOT three different models each with one predictor.**


```{r}
lm_multiple_regression <- lm(data=air, Ozone ~ Solar.R + Wind + Temp)
summary(lm_multiple_regression)
```

H.	Report the **adjusted R-Squared** in a comment – how does it compare to the adjusted R-squared from Step F? Is this better or worse? Which of the predictors are **statistically significant** in the model? In a comment, report the coefficient of each predictor that is statistically significant. Do not report the coefficients for predictors that are not significant.


```{r}
# The Adjusted R-squared is 0.4207 for this multiple regression model, as opposed to 0.2527 for the simple regression model.
# This is better because a higher adjusted R-squared value means we can account for more of the change in Ozone using
# multiple variables than we were able to account for using only 1 variable.
# Wind and Temp are statistically significant since their p-values are less than 0.05, but Solar.R is not statistically significant
# since its p-value is more than 0.05.
# The coefficient for Wind is -2.69669. This means that as Wind increases by 1, Ozone will decrease by -2.69669.
# The coefficient for Temp is 1.53072. This means that as Temp increases by 1, Ozone will increase by 1.53072.
```

I.	Create a one-row data frame like this: 


```{r}
predDF <- data.frame(Solar.R=290, Wind=13, Temp=61)
```

 and use it with the **predict( )** function to predict the **expected value of Ozone**:


```{r}
predict(lm_multiple_regression, predDF)
```

J.	Create an additional **multiple regression model**, with **Temp** as the **outcome variable**, and the other **3 variables** as the **predictors**. 

Review the quality of the model by commenting on its **adjusted R-Squared**.  


```{r}
# The Adjusted R-squared is 0.403, meaning that ~40% of the change in Temp can be explained by a change in Ozone,
# Solar.R, and Wind. All coefficients are statistically significant.
lm_multiple_regression_temp <- lm(data=air, Temp ~ Ozone + Solar.R + Wind)
summary(lm_multiple_regression_temp)
```
